# In-Class Programming Activity, Day 5

## Mira

#Task 1
```{r,include=FALSE}
library(mosaic)
nbins<-10
```

#Task 2
```{r}
evenly_spaced<-seq(min(height,data=Galton),max(height,data=Galton), length= (nbins+1))

```

#Task 3

cut(Galton$height, breaks=evenly_spaced) : finds where each case in height would fit into which interval from evenly_spaced


```{r}
bin_counts<-table(cut(Galton$height, breaks=evenly_spaced))


```


#Task 4
```{r}
xL<-evenly_spaced[1:(length(evenly_spaced)-1)]
xR<-evenly_spaced[2:length(evenly_spaced)]
count<-as.numeric(bin_counts)
hist_basics<-data.frame(xL,xR,count)
```


#Task 5
```{r}
make_one_bar <- function(point) {
  xLeft <- point$xL
  xRight <- point$xR
  height <- point$count
  res <- data.frame(
              x = c(xLeft,  xLeft,  xRight, xRight, NA),
              y = c(    0, height,  height,      0, NA)
  )
  res
}


one_to_five <- function(hist_data) {
  bars <- NULL
  for (k in 1:nrow(hist_data)) {
    new_bar <- make_one_bar(hist_data[k, ])
    bars <- rbind(bars, new_bar)
  }
  bars
}
My_bars<-(one_to_five(hist_basics))


```


Task 6
```{r}
plot(My_bars,type="n")
lines(My_bars)
polygon(My_bars,col="Blue")

```
Draw the histogram.

plot(My_bars, type="n") will make an empty plotting frame to hold the histogram.
lines(My_bars) will draw the bounding lines of each bar.
polygon(My_bars) will draw a polygon. The col = argument will set the color.


## Test statements

```{r}
scoreActivity::score253(5)
```