# In-Class Programming Activity, Day 7

# Mira Ensley-Field

# Likelihood Calculations

A likelihood is a probability of a set of observations given a model. In the simplest terms, a probability is a number between zero and one. In practice, one works with the logarithm of the likelihood, which will be a number between 0 and −∞−∞. Of course, at the point where the likelihood has a maximum, the log-likelihood will also have a maximum. The reason to prefer log-likelihoods is that likelihoods can be very, very small, beyond the capabilities of numerical calculations in computer hardware. The logarithm of the likelihood, however, is easily used in numerical calculations.

Create exponential function
```{r}
vals<-rexp(n=10,rate=1/100)
test200<-sum(log(dexp(x=vals,rate=1/200))) #gives log likelihood

LL_exp<-function(rate){  
  sum(log(dexp(x=vals,rate=rate)))
}

rates<-seq(1/200,1/50,length=100)
results<-sapply(rates,LL_exp)
plot(results~(1/rates))

optimize(f=LL_exp,lower=1/200,upper=1/50,maximum=TRUE)
exp_results<-optimize(LL_exp,lower=1/200,upper=1/50,maximum=TRUE)
```

## Test statements

```{r}
scoreActivity::score253(7)
```