# In-Class Programming Activity, Day 8

# Mira

#Task 1 — Fitting a line with likelihood


```{r}
x<-runif(100,20,70)
y<-5+3*x + 2*rnorm(100)
My_data<-data.frame(x,y)

plot(My_data)

LL_line<-function(params) {
  m<-params[1]
  b<-params[2]
  sigma<-params[3]
  sum(log(dnorm(y - (m * x + b), sd = sigma)))
}

testA<-LL_line(c(3,5,2))
testB<-LL_line(c(4,1,10))

starting_params<-c(4,1,10)

best<-optim(starting_params,LL_line,control=list(fnscale=-1))

print(best)
```

optim is a brute force approach, which is needed when the error isn't normally distrubted or changes with xx

#Task 2 Taxicab fare structure
-trying to model fare amount which depends on trip distance in reliable way
-depends  on waiting time, which lets assume has exponential distribution

  fare amount = trip distance * per mile + base fare + error
  error(waiting time)=fare amount-trip distance*permile - base fare

```{r}
load(url("http://tiny.cc/dcf/Taxi_trips.rda"))

starting<-c(base_fare = 2.00, per_mile = 4.00, params = 1/3)

taxi_likelihood<-function(vector){
  base_fare<-vector[1]
  per_mile<-vector[2]
  params<-vector[3]
  sum(log(rexp(Taxi_trips$fare_amount-base_fare-per_mile*Taxi_trips$trip_distance,params)))



}

best<-optim(par=starting,fn=taxi_likelihood,control=list(fnscale=-1)) #producing a lot of NAs
plot(Taxi_trips$trip_distance, Taxi_trips$fare_amount)
abline(best$base_fare,best$per_mile,col="red")
```

```{r}
plot(Taxi_trips$trip_distance, Taxi_trips$fare_amount)
abline(2.0329219, 4.7489711,col="red")
abline(2.3029117,3.6084402,col="blue")
abline(2.4019141, 3.8097856,col="green")
```

## Test statements

```{r}
scoreActivity::score253(8)
```