# In-Class Programming Activity, Day 9

# Mira Ensley-Field

Royal flush: <0.001% five cards of the same suit with ranks 10, ack, queen, king, ace. This could also be called a royal straight flush, but a royal flush is by necessity a straight flush.

Straight flush (not including royal flush): <0.002% five cards of the same suit in consecutive order


Flush (excluding royal flush and straight flush): 0.20%
five cards of the same suit

Straight (excluding royal flush and straight flush): 0.39%
five consecutively numbered cards


Deck, suit, and rank
```{r}
poker_deck <- c(outer((10 * 2:14), 1:4, '+'))
#suit: 1, 2, 3, 4
suits<-function(digits) {
  return(digits %% 10)
}

ranks<-function(digits) {
  return (digits %/% 10)
}
```

is_a functions
```{r}
#five cards of the same suit with ranks 10, ack, queen, king, ace.
is_royal_flush<-function(hand) {
    counts <- table(ranks(hand)) 
    counts1<- table(suits(hand))
    table(hand)==table(order(hand)) && min(counts)==10 && length(counts1)==1
}

#five cards of the same suit in consecutive order
is_straight_flush<-function(hand) {
  countSuit<-table(suits(hand))
  ranks(hand)==sort(ranks(hand),decreasing=FALSE) && length(countSuit)==1
}

is_four_of_a_kind<-function(hand){
  counts<-table(ranks(hand))
  max(counts)==4 && min(counts) == 1
  }

#Three of a kind and the other two cards forming a pair
is_full_house<-function(hand){
  counts <- table(ranks(hand))
  min(counts) == 2 && max(counts) == 3
}

#5 cards of same suit, not royal flush
is_flush<-function(hand) {
  counts <- table(suits(hand))
  length(suits)==1
}

#5 consectuively numbered cards (but not all of same suit)
is_straight<-function(hand) {
  countSuit<-table(suits(hand))
  ranks(hand)==sort(ranks(hand),decreasing=FALSE) && length(countSuit)>=1
}


is_three_of_a_kind<-function(hand){
  counts<-table(ranks(hand))
  max(counts)==3 && length(counts)==3
}

is_two_pair<-function(hand) {
    counts<-table(ranks(hand))
    max(counts)==2 && min(counts)==1 && length((counts))==3
}

is_pair<-function(hand) {
    counts<-table(ranks(hand))
    max(counts)==2 && min(counts)==1 && length((counts))==4
}
```

Testing is_a functions
```{r}

test_hand_2<-sample(poker_deck,5)

test_hand_1<-c(132, 142,  23,  33,  43)
testC<-c(132, 133,  134,  24,  41)
testD<-c(132, 133,  23,  24,  41)
testE<-c(132, 133,  23,  34,  41)
straight<-c(21, 31, 42, 52, 62)
sflush<-c(21, 31, 42, 52, 62)
sflushy<-c(21, 31, 41, 51, 61)

```

Estimating probabilities

Write a function before_draw() that draws five random cards from the deck and calculates the best hand that can be assembled from those cards. It will look like:

```{r}

before_draw <- function(poker_deck) {
  five_cards <- sample(poker_deck, 5,replace=TRUE)
  ordered_five_cards=sort(five_cards,descending=FALSE)
if(is_straight_flush(five_cards)) {return ("straight flush")}
if(is_four_of_a_kind(five_cards)) {return (" four of a kind")}
if(is_full_house(five_cards)) {return ("full house")}
if(is_flush(five_cards)) {return ("flush")}
if(is_straight(five_cards)) {return ("straight")}
if(is_three_of_a_kind(five_cards)) {return ("three of a kind")}
if(is_two_pair(five_cards)) {return ("two pair")}
if(is_pair(five_cards)) {return ("pair")}

}

```


```{r}
table(sapply(1:10000, FUN=before_draw))
```



## Test statements

```{r}
scoreActivity::score253(9)
```