# In-Class Programming Activity, Day 12

Mira Ensley-Field

# Testing LDA and QDA


```{r}
library(mosaic)
library(mosaicData)
library(MASS)
```


```{r}
n_cases<-1000
red_mean<-c(1,0)
green_mean<-c(0,-1)
blue_mean<-c(-1,1)
covar_red<-cbind(c(3,-1.7),c(-1.7,1))
covar_blue<-cbind(c(2,1.5),c(1.5,3))

One<-cbind(rnorm(n_cases,0,1),rnorm(n_cases,0,1))
Two<-cbind(rnorm(n_cases,0,1),rnorm(n_cases,0,1))
Three<-cbind(rnorm(n_cases,0,1),rnorm(n_cases,0,1))
```

Creating then modifying RGB Matrices
```{r}
Red<-One %*% chol(covar_red)
Green<- Two %*% chol(covar_red) 
Blue<- Three %*% chol(covar_blue)

Red<-cbind(Red,matrix(red_mean,nrow=1000))
Green<-cbind(Green,matrix(green_mean,nrow=1000))
Blue<-cbind(Blue,matrix(blue_mean,nrow=1000))

Red[1]<-Red[1] +1
Green[1]<-Green[1]+1
Blue[1]<-Blue[1]+1


```

Creating three dataframes
```{r}
Red<-data.frame(Red[,1],Red[,2],class="red")
names(Red)<-c("x","y","class")
Blue<-data.frame(Blue[,1],Blue[,2],class="blue")
names(Blue)<-c("x","y","class")
Green<-data.frame(Green[,1],Green[,2],class="green")
names(Green)<-c("x","y","class")

```

Creating final 2 simulated dataframes
```{r}
Same_sigma<-rbind(Red,Green)
Different_Sigma<-rbind(Red,Blue)
```


LDA and Confusion Matrix 1
```{r}
mod_LDA_one <- lda(class~x+y, data = Same_sigma)
test_LDA_one <- predict(mod_LDA_one, newdata = Same_sigma)
table(Same_sigma$class, test_LDA_one$class)
```
QDA and Confusion MAtrix 2
```{r}
mod_QDA_one <- qda(class~x+y, data = Same_sigma)
test_QDA_one <- predict(mod_QDA_one, newdata = Same_sigma)
table(Same_sigma$class, test_QDA_one$class)
```

Compare LDA and QDA 1
```{r}
lda_test_error<-mean(Same_sigma$class != test_LDA_one$class)
qda_test_error<-mean(Same_sigma$class != test_QDA_one$class)
print(lda_test_error) 
print(qda_test_error) 
```
small n: the qda has a  lower test error rate (.43), though they are both relatively high (.355)
big n: They are virtually the same

LDA 2
```{r}
mod_LDA_two <- lda(class~x+y, data = Different_Sigma)
test_LDA_two <- predict(mod_LDA_two, newdata = Different_Sigma)
table(Different_Sigma$class, test_LDA_two$class)
```
QDA 2
```{r}
mod_QDA_two <- qda(class~x+y, data = Different_Sigma)
test_QDA_two <- predict(mod_QDA_two, newdata = Different_Sigma)
table(Different_Sigma$class, test_QDA_two$class)
```

Compare LDA and QDA 2
```{r}
lda_test_error<-mean(Different_Sigma$class != test_LDA_two$class)
qda_test_error<-mean(Different_Sigma$class != test_QDA_two$class)
print(lda_test_error)  
print(qda_test_error) 
```
small n: The error rate on the LDA (.465) is much lower than the QDA (.06)
big n: The error rate on the LDA (.481) is much lower than the QDA (.0885)

##Bigger N



## Test statements

```{r}
scoreActivity::score253(12)
```