# In-Class Programming Activity, Day 14

Your work goes here ...

#Mira

#Shrinkage and out-of-sample prediction error

https://dtkaplan.github.io/math253/ProgrammingActivities/Day-14-Programming-Task.html

```{r}
library(glmnet)
GenCont <- read.csv("http://tiny.cc/dcf/GenCont.csv")
```


pred_lda <- predict(lda_model, newdata = test_data)
table(pred_lda$class, test_data$mpg01)
mean(pred_lda$class != test_data$mpg01)

lasso_mod_real<-cv.glmnet(x=X, y=Y_real, alpha=1)
beta_lasso_real <- predict(lasso_mod, type = "coefficients", s = lasso_mod$lambda.min)
plot(as.numeric(beta_lasso_real)[-1],beta)

Comparing OLS to ridge
```{r}
compare_ols_ridge<-function(responses=GenCont[,1],predictors=GenCont[,-1],lambda=1) {

 #training and testing data
  train_rows<-sample(1:length(responses), size = floor(length(responses)/2))

  GenCont_train<-GenCont[train_rows,]
  GenCont_test<-GenCont[-train_rows,]
  responses<-as.matrix(GenCont[,1])
  predictors<-as.matrix(GenCont[,2:13])
  
  #predictors
  train_data_pred <-as.matrix( predictors[train_rows,])
  test_data_pred <- as.matrix(predictors[-train_rows, ])
  
  #response
  train_data_resp <- as.matrix(responses[train_rows,])
  test_data_resp <- as.matrix(responses[-train_rows, ])


  #OLS (Linear Regression)
  ols_mod<-lm(train_data_resp~train_data_pred)
  ols_in_pred<-predict(ols_mod)
  ols_out_pred<-predict(ols_mod,newx=test_data_resp)
  
  #MSPE
  ols_in<-mean(ols_in_pred-train_data_resp)^2
  ols_out<-mean(ols_out_pred-test_data_resp)^2
  
  #glmnet (Ridge Regression) GLM_MODnot working
  
  #glm_mod<-glmnet(x=train_data_pred,y=train_data_resp,alpha=0,lambda=lambda)
  #glm_pred<-predict(glm_mod,newx=train_data_resp)
  
  #MSPE
  ridge_in<-mean((ols_in_pred - train_data_pred)^2)
  ridge_out<-mean((ols_out_pred-test_data_resp)^2)

  Shrink_results<-data.frame(lambda,ols_in,ols_out,ridge_in,ridge_out)

  return(Shrink_results)
}
```

```{r}
compare_ols_ridge(predictors=GenCont[,2:13],responses=GenCont$Phenotypes)
```

Exploring Lambda
```{r}
lambda_long<-seq(.1,100,1)
Shrink_results<-compare_ols_ridge(predictors=GenCont[,2:13],responses=GenCont$Phenotypes,lambda=lambda_long)
```
  
  
## Test statements

Your function should return a named vector with these five components:
lambda as well as ols_in, ridge_in, ols_out, ridge_out,
each of which has the corresponding mean square prediction error.
8. Use your function to create an object called Shrink_results with
lambda = 1 and the GenCont data .
Exploring Lambda
Use your compare_ols_ridge() to look at the in- and out-of-sample
prediction errors for a range of λ from 0.1 to 100


```{r}
scoreActivity::score253(14)
```

