# In-Class Programming Activity, Day 16

Your work goes here ...

# Mira
# Knots, cubic spline fitter

```{r}
library(ISLR)
```

the knots: returns k quariles with values 1/(k+1), 2/(k+1)...etc 
probs: for #'s 1 thru k, calculates the prob
```{r}
z<-runif(100,0,10)

my_knots<-function(x=z,k=5){
  quantile(x,probs=(1:k)/(k+1))
}
```


```{r}
my_knots(z,5)
```

The basis set
creates a cubic model matrix: first four are 'built in' functions, any additional must be specified
```{r}
spline_model_matrix<-function(x,knots){
 MM<-cbind(1,x,x^2,x^3)
  for (i in 1:length(knots)) {
    new_vector<-ifelse(x<knots[i],0,(x-knots[i])^3)
    MM<-cbind(MM,new_vector)
  }
  MM
 }
```

# creates three new_vectors to help approximate z
```{r}
test<-spline_model_matrix(z,my_knots(z,3))
```

 You’ll have to go back to
fit_spline() and arrange for it to include the rse. This is the sum of
square residuals divided by the degrees of freedom in the residual.3

Finding best linear combination
```{r}
fit_spline <- function(formula, k=2, data=parent.frame()) {
  y <- eval(formula[[2]], envir=data)
  x <- eval(formula[[3]], envir=data)
  knot_locations <- my_knots(x, k)
  MM <- spline_model_matrix(x, knot_locations)
  mod <- lm(y ~ MM - 1)
  res <- list(coef = coef(mod), knots = knot_locations, cov = vcov(mod))
  class(res) <- "my_spline"
  return(res)
}

```

```{r}
mod<-fit_spline(data=mtcars,k=2,formula=mpg~hp)
```

# My Spline (simple)

```{r}
predict.my_spline <-function(mod, newx, level = 0.95, intervals=c("none", "confidence", "prediction")) {
  intervals <- match.arg(intervals)
  MM <- spline_model_matrix(newx, mod$knots)
  vals <- MM %*% mod$coef
  se <- sqrt(rowSums(MM %*% mod$cov * MM))
  
  if (intervals == "none") return(vals)
  else return(NULL) # for future use
}
```


```{r}
mod<-fit_spline(data=Wage,k=17,formula=wage~age)

plot(wage~age,data=Wage)
x<-0:100 
y<-predict.my_spline(mod, newx=x)
lines(x,y,col="red")
```

Confidence intervals
```{r}
predict.my_spline <-function(mod, newx, level = 0.95, intervals=c("none", "confidence", "prediction")) {

  #intervals <- match.arg(intervals)

  MM <- spline_model_matrix(newx, mod$knots)
  se <- sqrt(rowSums(MM %*% mod$cov * MM))
  vals <- MM %*% mod$coef
  upper<-vals + 2*se
  lower<-vals-2*se
  conf<-data.frame(vals,upper,lower)
  
   if(intervals=="confidence") {
     print("CONFIDENCE")
    return(conf)
  }  else return(NULL) # for future use
}
```



```{r}
mod<-fit_spline(data=Wage,k=17,formula=wage~age)
plot(wage~age,data=Wage)
x2<-1:100
y2<-predict.my_spline(mod, newx=x2,intervals="confidence")
lines(x2,y2[,1],col="red") 
lines(x2,y2[,2],col="green") 
lines(x2,y2[,3],col="blue") 


```

will return the three-column data frame,
but giving the 95% prediction intervals. These will be based on se
and the standard error of the residuals rse from the fit. The prediction
standard error is sqrt(seˆ2 + rseˆ2).

Prediction intervals

```{r}
fit_spline <- function(formula, k=2, data=parent.frame()) {
  y <- eval(formula[[2]], envir=data)
  x <- eval(formula[[3]], envir=data)
  knot_locations <- my_knots(x, k)
  MM <- spline_model_matrix(x, knot_locations)
  mod <- lm(y ~ MM - 1)
  res <- list(coef = coef(mod), knots = knot_locations, cov = vcov(mod))
  rse<-sum(coef(mod))
  class(res) <- "my_spline"
  return(res)
}

```


```{r}
predict.my_spline <-function(mod, newx, level = 0.95, intervals=c("none", "confidence", "prediction")) {

  #intervals <- match.arg(intervals)

  MM <- spline_model_matrix(newx, mod$knots)
  se <- sqrt(rowSums(MM %*% mod$cov * MM))
  vals <- MM %*% mod$coef
  upper<-vals+2*sqrt(se*se + rse*rse)
  lower<-vals-2*sqrt(se*se + rse*rse)

  conf<-data.frame(vals,upper,lower)
  
   if(intervals=="prediction") {
     print("PREDICTION")
    return(conf)
  }  else return(NULL) # for future use
}
```


```{r,eval=FALSE}
mod<-fit_spline(data=Wage,k=17,formula=wage~age)
#plot(wage~age,data=Wage)
x3<-1:100
#y3<-predict.my_spline(mod, newx=x3,intervals="prediction")
#lines(x2,y2[,1],col="red") 
#lines(x2,y2[,2],col="green") 
#lines(x2,y2[,3],col="blue") 


```


## Test statements

```{r}
scoreActivity::score253(16)
```